---
title: "jbsMisc"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

```{r, echo = FALSE}
library(tidyverse)
library(ggplot2)
```


## Installation


Run the command:

```
remotes::install_github("jb-santos/svyEffects")
```

You'll need the `{remotes}` package to do so, which can be installed by running:

```
install.packages("remotes")
```


## Simple crosstab using `ctab()`


Create a survey design object. (These functions also work on vanilla data frames).

```{r}
library(tidyverse)
library(survey)
data("ab")
absvy <- svydesign(~1, data = ab, weights = ~weight)
```

Generate the crosstab using `ctab()`.

```{r}
library(jbsMisc)
table_vote <- ctab(absvy, yvar = "dvote3", xvar = "region")
table_vote
```

The output is in long, or "tidy" format and lends itself well to plotting with `ggplot()`.

```{r}
library(ggplot2)
ggplot(table_vote) +
  aes(x = region, y = pct, fill = dvote3) +
  geom_bar(stat = "identity")
```


```{r}
library(ggplot2)
ggplot(table_vote) +
  aes(x = region, y = pct, fill = dvote3) +
  geom_bar(stat = "identity", position = position_dodge())
```


## Multiple crosstabs (aka "stub-and-banner" tables) using `ctabs()`


We can generate stub-and-banner crosstabs that cross-tabulate a single "dependent" (or row) variable against multiple "independent" (columns) variables using the `ctabs()` function. This is similar to the output in many polling reports (the generation of which is what led me to write these functions).

Calculate the crosstabs. Multiple column variables can be specified by naming multiple variables within the argument `xvars = c()`.

The output is a list object with the class `ctabs`. 

```{r}
library(jbsMisc)
table_vote1 <- ctabs(absvy, yvar = "dvote3", xvars = c("region", "female", "age"))
names(table_vote1)
```

The first slot `$Summary` is the multiple crosstab table. 

```{r}
table_vote1
```

The second slot `$Total` contains the marginal frequency distribution of the row variable. 

```{r}
table_vote1[[2]]
```

Subsequent tables contain a long- or tidy-format crosstab between the row variable and the respective column variable.

```{r}
table_vote1[[4]]
```


### Plotting multiple crosstabs


There is a `plot()` method included for the objects outputted by `ctabs()`.

```{r}
plot(table_vote1)
```

The `plot()` function can take a few arguments to modify the default plots.

`plot.ctabs(x, tblno = 1, dodge = FALSE, txt = FALSE)`

- `tblno = ` selects which table in the `ctabs` list object to plot.
- `dodge = ` whether bars should be "clustered" using `position_dodge()`. By default, they are stacked using `position_stack()`.
- `txt = ` should text labels be added to graph.


```{r}
plot(table_vote1, dodge = TRUE, txt = TRUE)
```

`plot()` is a wrapper for `ggplot()`, so you can add additional options using `+ ...`.

```{r}
plot(table_vote1, dodge = FALSE, txt = TRUE) +
  scale_fill_manual(values = c("#184484", "#f37221", "#bbbbbb")) +
  labs(title = "Alberta vote intention",
       subtitle = "Decided/leaning voters only, n=1,074",
       y = "",
       x = "Sub-group",
       fill = "Party",
       caption = "Janet Brown Opinion Research / Trend Research") +
  theme_classic()
```


### Plotting other tables in `ctabs()` objects


```{r}
plot(table_vote1, 2, dodge = TRUE) + theme_bw()
```

```{r}
plot(table_vote1, 5, dodge = TRUE) + theme_bw()
```


## Testing columns for significant differences using `testcols()`


This function tests columns for significant differences. You can also obtain Holm-corrected p-values for multiple comparisons using the argument `adj.p = TRUE`.

For now, it outputs test result summaries (i.e. significance stars) to a separate table. I'll eventually update the output when I find some time in the future.

```{r}
table_vote1 <- ctabs(absvy, yvar = "dvote3", xvars = c("region", "female", "age"))
table_vote1_tests <- testcols(table_vote1)
```


```{r}
table_vote1_tests
```

```{r}
table_vote1[[3]] %>% 
  select(-n) %>%
  pivot_wider(names_from = region, values_from = pct)
table_vote1_tests[[1]]
```


```{r}
table_vote1[[4]] %>% 
  select(-n) %>%
  pivot_wider(names_from = female, values_from = pct)
table_vote1_tests[[2]]
```

```{r}
table_vote1[[5]] %>% 
  select(-n) %>%
  pivot_wider(names_from = age, values_from = pct)
table_vote1_tests[[3]]
```

